Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-03-16 paper/current/main.tex
changed clustering fig text

Just as the embedding space makes it possible to \textit{zoom in} on subsets of organizations, it is also possible to \textit{zoom out} by aggregating organizational vectors.
In doing so, we can examine the country-level structure that governs scientific mobility.
We define the representative vector of each country as the average of their organizational vectors and, using their cosine similarities, perform hierarchical clustering of nations that have at least 25 organizations represented in the embedding space, excluding the United States which is a dominant hub well connected to most other countries (see Fig. \ref{fig:clustering}a).
The six identified clusters roughly correspond to countries in East-Asia (orange), Scandinavia (dark purple), the British Commonwealth (light purple), and Central and Eastern Europe (light blue), and then two remaining diverse clusters contain a mix of European, Latin American, and Mediterranean countries.
The cluster structure shows that not only geography but also linguistic ties between countries are related to scientific mobility.

@@

%
% Figure - CLUSTERING
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Clustering/cluster_country.pdf}
	\caption{
		\textbf{Geography, then language, conditions international mobility.}
		\textbf{a.} Hierarchically clustered similarity matrix of country vectors aggregated as the mean of all organization vectors within the country, excluding the United States and countries with less than 25 organizations.
		Color of matrix cells corresponds to the cosine similarity between country vectors.
		Color of country names corresponds to their cluster.
		Color of three cell columns separated from the matrix corresponds to, from left to right, the region of the country, the language family~\autocite{ethnologue}, and the dominant language.
		\textbf{b.}	Element-centric cluster similarity~\autocite{gates2019element} reveals the factors dictating hierarchical clustering.
		Region better explains the grouping of country vectors at higher levels of the clustering.
		Language family, and then the most widely-spoken language, better explain the fine-grained grouping of countries.
	}
	\label{fig:clustering}
\end{figure}

>>>>

Just as the embedding space makes it possible to \textit{zoom in} on subsets of organizations, it is also possible to \textit{zoom out} by aggregating organizational vectors.
In doing so, we can examine the country-level structure that governs scientific mobility.
We define the representative vector of each country as the average of their organizational vectors and, using their cosine similarities, perform hierarchical clustering of nations that have at least 25 organizations represented in the embedding space (see Fig. \ref{fig:clustering}a).
The six identified clusters roughly correspond to countries in Asia and North America (orange), Northern Europe (dark blue), the British Commonwealth and Iran (purple), Central and Eastern Europe (light blue), South America and Iberia (dark green), and Western Europe and the Mediterranean (light green). 
The cluster structure shows that not only geography but also linguistic ties and cultural between countries are related to scientific mobility.

@@

%
% Figure - CLUSTERING
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Clustering/cluster_country.pdf}
	\caption{
		\textbf{Geography, then language, conditions international mobility.}
		\textbf{a.} Hierarchically clustered similarity matrix of country vectors aggregated as the mean of all organization vectors within countries with at least 25 organizations.
		Color of matrix cells corresponds to the cosine similarity between country vectors.
		Color of country names corresponds to their cluster.
		Color of three cell columns separated from the matrix corresponds to, from left to right, the region of the country, the language family~\autocite{ethnologue}, and the dominant language.
		\textbf{b.}	Element-centric cluster similarity~\autocite{gates2019element} reveals the factors dictating hierarchical clustering.
		Region better explains the grouping of country vectors at higher levels of the clustering.
		Language family, and then the most widely-spoken language, better explain the fine-grained grouping of countries.
	}
	\label{fig:clustering}
\end{figure}
